Should I Airbnb My Property?

### 2017-03-08 abb_Writeup.Rmd
results - metro and type

As the observed calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the Actual revenue metric to be a better measure of actual market performance. We chart the distribution of the Actual ($REV_a$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long term revenues for the long term properties. Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year (Figure 8).  Actual short-term revenues range from neglible to highs of greater than \$80,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is much flatter, with many more properties showing potential incomes in the \$20,000 to \$40,000 range.

@@

```{r plot_revdens, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5, fig.cap='Comparison of Revenue Distributions', fig.lp='revdens'}

  stre <- str.data[,c('bedrooms', 'str.ext.revenue')]
  stre$tenure <- 'short-term extrapolated   '
  names(stre)[2] <- 'revenue'
  strp <- str.data[,c('bedrooms', 'str.pot.revenue')]
  strp$tenure <- 'short-term potential   '
  names(strp)[2] <- 'revenue'
  ltrr <- ltr.data[,c('bedrooms', 'ltr.revenue')]
  ltrr$tenure <- 'long-term   '
  names(ltrr)[2] <- 'revenue'
  allr <- rbind(stre, strp, ltrr)

@@

Occupancy rates are one of the most important components in determining the profitability of short-term rentals. Below we show (left-panel, Figure 9) the percentage of properties for which short-term rental is more profitable versus the occupancy rate over the year analyzed. We see that likelihood of short-term rental being a better option increases sharply with occupancy rate up from about 25\% up to 75\% occupancy. For occupancy rates less than 40\%, long-term rental is more likely to be preferable.

As the distribution of occupancy rates is not uniform the initial plot may be misleading as to the frequency of short-term rentals being more profitable than long term. In the right hand panel of Figure 9 we transform the x-axis to represent the quantile of the full distribution of occupancy rates. By doing so, we see that, in general, properties need to be in the upper 25th percentile of occupancy rates across the entire market in order to have a equal likelihood of being more profitable as a short-term rental than as a more traditional long term one. As a matter of reference, the mean occupancy rate in the sample is 27.6\%, with a median value of 19.4\%.

@@

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=7, fig.height=3.5, fig.cap='Probability of Short Term Pref. vs Occ. Rate'}

@@

Along with occupancy rates, the other important consideration is the nightly rate.  High occupancy rates with low nightly rates may not be any better than lower occupancy rate with a higher nightly rate.  In Figure 10 (left panel) we plot nightly rates versus occupancy rates.  We then divide the area of the scatter plot into 400 equal areas (20 x 20) and color each based on which is more common in the area, short term or long term revenue.  The transparency is related to the total count of actual market properties falling into that area.

The general trend is what is expected, i.e. that higher rates and occupancies, ceteris paribus, means more profitability for short term rentals.  This relationship is not a perfectly linear because the long-term rental revenue varies greatly based on the characteristics of the property.  While the left panel of Figure 16 helps show the general pattern is does leave many indeterminate areas.  In order to better delimit the situations where short-term rentals are more profitable we turn to a support vector machine (SVM) algorithm.

An SVM can be used as a classifier to delineate the boundaries between two (or more) classes of observations.  In this case, we delineate the boundary (right panel) between situations where long-term rentals are more profitable (red) and where short term are (green).  In order to avoid an overly broad straight line delineation we use a polynomial fitting technique. While this process does provide more flexible fit in the area of high data density, it also has the side effect of creating somewhat counter intuitive results in areas of low data (such as the far upper right corner.)  Overall, from the SVM model we see that the short-term preference begins at about 30\% occupancy rate, regardless of nightly rate.  As the occupancy rate increases, the required nighly rate decreases, down to near $100 once occupancies hit 65\% or greater.

@@

```{r, fig.width=8, fig.height=4, echo=FALSE, fig.align='center', fig.cap='Analysis of Raw Rates: Metro Level'}

@@

Both the occupancy rates and the nightly rates are not uniformly distributed and therefore the plots in Figure 10 may be misleading.  We convert both axis to quantiles and replot both in Figure 11.  Here we see a more even distibution of points across the figures.  Additionally, a more reasonable SVM breakdown occurs as well.  The right panel shows that in order for the likelihood of a property to be more profitable in short-term rental to be above 50\% the lowest occupancy rate it can have is 40\%, if the nightly rate is in the top 95th percentile of the market.  For example, if we drop the nightly rate to the median, the occupancy rate need to be at the 60th percentile.

@@

  qtl.hm <- abbHeatMap(str.data.ic,
                x.field='occ.qtl',
                y.field='rate.qtl',
                pref.field='str.ext.pref',
                svm=F, 
                alpha.count=T,
                add.points=T,
                fill.colors=c(abb.col[1], abb.col[5]))
  qtl.hm <- qtl.hm + ylab('Nightly Rate')

  qtl.hm.svm <- abbHeatMap(str.data.ic,
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.ext.pref',
                   svm=T, 
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))
  qtl.hm.svm <- qtl.hm.svm + ylab('Nightly Rate')

@@

To gain a deeper understanding of the differences in short-term preference, we break the data down across a variety of characteristics and examine the preferability metrics in these submarkets. First, we bifurcate our data into houses and apartments.  Within the houses, 17.5\% of the properties are short-term preferable, while for the apartment that figure is nearly double at 30.5\%. Plotting the changes in short-term preferability against occupancy rates shows considerable similarity between houses and apartments (Figure 12).  The only measurable difference is in that for houses a properties occupancy rate needs to be about in a higher quantile (by about 10) than apartments to show the same likelihood of short-term preference as apartment.  Or, in other words, a given house needs to outperform its peers by a greater extent than an apartment to achieve short term preferenc.e

@@

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=7, fig.height=3.5, fig.cap='Preferency vs Occ. Rates: By Type'}

@@

  ptq.occ <- ggplot(ptqocc$data,
                    aes(x=x.var, y=Var, group=ID, color=ID)) +  
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate (Quantile)') +
    scale_color_manual(values=c('black', 'grey60'),
                       name='') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25th', '50th', '75th', '100th')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom') + 
    ggtitle('Quantile Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

@@

Extending this analysis to also include nightly rates shows a similar situation (Figure 13). For houses (left panel) there are very few combinations of occupancy rate and nightly rate that offer a greater than even likelihood of short-term preference.  As an example, a median nightly rate requires a 90\% occupancy rate. The market for apartment (right panel) is more forgiving and occupancy rates as low as the 60th percentile can equate to a greater than even chance of short-term prefence is coupled with nightly rates near the top of the market.  A median nightly rate requires a 70\% occupancy rate for short-term prefernce likelihood.

@@

```{r, fig.width=8, fig.height=4, echo=FALSE, fig.cap='Analysis of Quantile Rates: By Type'}

@@

  house.hm.svm <- abbHeatMap(type.data.ic[type.data.ic$type ==
                                          'House', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.ext.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5])) +
  theme(legend.position='none') +
  annotate("text", x = 50, y = 50, label = "Houses", size=9)

  apt.hm.svm <- abbHeatMap(type.data.ic[type.data.ic$type ==
                                          'Apartment', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.ext.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))+
  theme(legend.position='none')+
  annotate("text", x = 50, y = 50, label = "Apartments", size=9)

>>>>

As the observed calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the Actual revenue metric to be a better measure of actual market performance. We chart the distribution of the Actual ($REV_a$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long term revenues for the long term properties. Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year (\autoref{revdens}).  Actual short-term revenues range from neglible to highs of greater than \$80,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is much flatter, with many more properties showing potential incomes in the \$20,000 to \$40,000 range.

@@

```{r plot_revdens, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5, fig.cap='\\label{revdens}Comparison of Revenue Distributions', fig.lp='revdens'}

  stre <- str.data[,c('bedrooms', 'str.ext.revenue')]
  stre$tenure <- 'short-term actual   '
  names(stre)[2] <- 'revenue'
  strp <- str.data[,c('bedrooms', 'str.pot.revenue')]
  strp$tenure <- 'short-term potential   '
  names(strp)[2] <- 'revenue'
  ltrr <- ltr.data[,c('bedrooms', 'ltr.revenue')]
  ltrr$tenure <- 'long-term   '
  names(ltrr)[2] <- 'revenue'
  allr <- rbind(stre, strp, ltrr)

@@

Occupancy rates are one of the most important components in determining the profitability of short-term rentals. For reference, the mean occupancy rate in the sample is 27.6\%, with a median value of 19.4\%. In \autoref{occpref} (left panel) we show the percentage of properties for which short-term rental is more profitable versus the occupancy rate over the year analyzed. We see that likelihood of short-term rental being a better option increases sharply with occupancy rate up from about 25\% up to 75\% occupancy. For occupancy rates less than 40\%, long-term rental is more likely to be preferable.

As the distribution of occupancy rates is not uniform the initial plot may be misleading as to the frequency of short-term rentals being more profitable than long term. In the right hand panel of \autoref{occpref} (right panel) we transform the x-axis to represent the quantile of the full distribution of occupancy rates. By doing so, we see that, in general, properties need to be in the upper 25th percentile of occupancy rates across the entire market in order to have a equal likelihood of being more profitable as a short-term rental than as a more traditional long term one.

@@

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=7, fig.height=3.5, fig.cap='\\label{occpref}Probability of Short Term Pref. vs Occ. Rate'}

@@

Along with occupancy rates, the other important consideration is the nightly rate.  High occupancy rates with low nightly rates may not be any better than lower occupancy rate with a higher nightly rate.  In Figure 10 (left panel) we plot nightly rates versus occupancy rates.  We then divide the area of the scatter plot into 400 equal areas (20 x 20) and color each based on which is more common in the area, short-term or long-term revenue.  The transparency is related to the total count of actual market properties falling into that area.

The general trend is what is expected; that is that higher rates and occupancies, ceteris paribus, means greater profitability for short-term rentals.  This relationship is not a perfectly linear because the long-term rental revenue varies greatly based on the characteristics of the property.  While the left panel of \autoref{metrohm} helps show the general pattern it does leave many indeterminate areas.  In order to better delimit the situations where short-term rentals are more profitable we turn to a support vector machine (SVM) algorithm.

An SVM can be used as a classifier to delineate the boundaries between two (or more) classes of observations.  In this case, we delineate the boundary, \autoref{metrohm} (right panel) between situations where long-term rentals are more profitable (red) and where short term are (green).  In order to avoid an overly broad straight line delineation we use a polynomial fitting technique. Overall, from the SVM model we see that the short-term preference begins at about 325% occupancy rate, regardless of nightly rate.  As the occupancy rate increases, the required nighly rate decreases, down to near $100 once occupancies hit 65\% or greater.

@@

```{r, fig.width=8, fig.height=4, echo=FALSE, fig.align='center', fig.cap='\\label{metrohm}Analysis of Raw Rates: Metro Level'}

@@

Both the occupancy rates and the nightly rates are not uniformly distributed and therefore the plots in \autoref{metrohm} may be misleading.  We convert both axis to percentiles and replot both in \autoref{metrohmp}.  Here we see a more even distibution of points across the figures.  Additionally, a more reasonable SVM breakdown occurs as well.  The right panel shows that in order for the likelihood of a property to be more profitable in short-term rental to be above 50\% the lowest occupancy rate it can have at around the 60th percentile of the market it is in.  And that only holds if the nightly rate is in the top 95th percentile of the market.  For example, if we drop the nightly rate to the median, the occupancy rate needs to be at the 75th percentile or above.

@@

```{r, fig.width=8, fig.height=4, echo=FALSE, fig.cap='\\label{metrohmp}Analysis of Percentile Rates: Metro Level'}

  qtl.hm <- abbHeatMap(str.data.ic,
                x.field='occ.qtl',
                y.field='rate.qtl',
                pref.field='str.ext.pref',
                svm=F, 
                alpha.count=T,
                add.points=T,
                fill.colors=c(abb.col[1], abb.col[5]))
  qtl.hm <- qtl.hm + ylab('Nightly Rate (Pcntl)') + xlab('Occupancy Rate (Pcntl)')

  qtl.hm.svm <- abbHeatMap(str.data.ic,
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.ext.pref',
                   svm=T, 
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))
  qtl.hm.svm <- qtl.hm.svm + ylab('Nightly Rate (Pcntl)') + xlab('Occupancy Rate (Pcntl)')

@@

To gain a deeper understanding of the differences in short-term preference, we break the data down across a variety of characteristics and examine the preferability metrics in these submarkets. First, we bifurcate our data into houses and apartments.  Within the houses, 17.5\% of the properties are short-term preferable, while for the apartment that figure is nearly double at 30.5\%. Plotting the changes in short-term preferability against occupancy rates shows considerable similarity between houses and apartments (\autoref{smocc}).  The only measurable difference is in that for houses a property's occupancy rate needs to be 10 percentiles higher than apartments to show the same likelihood of short-term preference as an apartment. Or, in other words, a given house needs to outperform its peers by a greater extent than an apartment to achieve short term preference.

@@

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=7, fig.height=3.5, fig.cap='\\label{smocc}Preference vs Occ. Rates: By Type'}

@@

  ptq.occ <- ggplot(ptqocc$data,
                    aes(x=x.var, y=Var, group=ID, color=ID)) +  
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate (Pcntl)') +
    scale_color_manual(values=c('black', 'grey60'),
                       name='') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25th', '50th', '75th', '100th')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom') + 
    ggtitle('Percentile Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

@@

Extending this analysis to include nightly rates shows a similar situation (\autoref{smhm}). For houses (left panel) there are very few combinations of occupancy rate and nightly rate that offer a greater than even likelihood of short-term preference.  As an example, a median nightly rate requires a 90th percentile occupancy rate. The market for apartment (right panel) is more forgiving and occupancy rates as low as the 60th percentile can equate to a greater than even chance of short-term prefence is coupled with nightly rates near the top of the market.  A median nightly rate requires a 70th percentile occupancy rate for short-term preference likelihood.

@@

```{r, fig.width=8, fig.height=4, echo=FALSE, fig.cap='\\label{smhm}Analysis of Percentile Rates: By Type'}

@@

  house.hm.svm <- abbHeatMap(type.data.ic[type.data.ic$type ==
                                          'House', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.ext.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5])) +
  theme(legend.position='none') +
  annotate("text", x = 50, y = 50, label = "Houses", size=9) +
  ylab('Nightly Rate (Pcntl)') +
  xlab('Occupancy Rate (Pcntl)')

  apt.hm.svm <- abbHeatMap(type.data.ic[type.data.ic$type ==
                                          'Apartment', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.ext.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))+
  theme(legend.position='none')+
  annotate("text", x = 50, y = 50, label = "Apartments", size=9)+
  ylab('Nightly Rate (Pcntl)') +
  xlab('Occupancy Rate (Pcntl)')
